All watched over by machines of loving grace

### 2021-08-02 paper/paper.tex
paper: minor changes to paper

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[colorlinks]{hyperref}
\usepackage{listings}
\usepackage{microtype}
\usepackage{prooftree}

@@

New types are ``carved out'' of non-empty subsets of existing types.
Given an existing type $\tau$ and a predicate $p : \tau \Rightarrow \mathtt{bool}$ we introduce a new type $\tau'$ whenever $\{\} \vdash \xsts{x_\tau}p(x_\tau)$ holds.
In addition, two new constants are entered into the global theory---$\mathtt{in}_{\tau'}$ with type $\tau \Rightarrow \tau'$ and $\mathtt{out}_{\tau'}$ with type $\tau' \Rightarrow \tau$, respectively---representing \deffont{injection} and \deffont{projection} functions for the new type.
These two functions are related by also entering $\{\} \vdash \fall{x_\tau}p(x_\tau) \longrightarrow out_{\tau'}(in_{\tau'}(x_\tau)) = x_\tau$ and $\{\} \vdash \fall{x_{\tau'}}in_{\tau'}(out_{\tau'}(x_{\tau'})) = x_{\tau'}$ as additional axioms.

@@

The Supervisionary kernel maintains a mutable runtime state consisting of a series of mutable \deffont{heaps}.
These heaps are indexed by \deffont{kernel handles}---unsigned integers generated by the kernel---and which can be thought of as pointers to \deffont{kernel objects} in the various heaps.
Importantly, the kernel never exposes kernel objects to untrusted code directly, but performs all manipulations on these objects itself, under instruction by untrusted code.

@@

The simplest---and most foundational---of these additional heaps is the \deffont{heap of type-formers}, which records the arities of type-formers registered with the kernel.
When untrusted code requests that a new type-former is registered with the kernel, it calls an ABI function---\texttt{TypeFormer.Handle.Register}---with a supplied arity.
In response, the kernel invents a fresh handle and associates it with the supplied arity, before returning the new handle back to untrusted code to use to refer to the type-former again in future.
Note that we can describe the action of the various ABI functions exposed by the kernel with Hoare triples, using an informal Separation Logic as the assertion language.

@@

\begin{figure}[t]
\caption{The internal structure of the kernel's type heap}
\label{fig.type.heap}
\end{figure}

@@

>>>>

\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[colorlinks]{hyperref}
\usepackage{listings}
\usepackage{microtype}
\usepackage{tikz}
\usepackage{prooftree}

\usetikzlibrary{matrix,backgrounds}

@@

A new type $\tau$ is ``carved out'' of an existing type $\tau'$ by identifying the non-empty collection of elements of $\tau'$ that belong to $\tau$.
Specifically, given a predicate $p : \tau \Rightarrow \mathtt{bool}$ we introduce a new type $\tau'$ whenever we can show $\{\} \vdash \xsts{x_\tau}p(x_\tau)$ holds.
In addition, two new constants are also entered into the global theory---$\mathtt{in}_{\tau'}$ with type $\tau \Rightarrow \tau'$ and $\mathtt{out}_{\tau'}$ with type $\tau' \Rightarrow \tau$, respectively---representing \deffont{injection} and \deffont{projection} functions for the new type.
These two functions are related by a pair of additional axioms, $\{\} \vdash \fall{x_\tau}p(x_\tau) \longrightarrow out_{\tau'}(in_{\tau'}(x_\tau)) = x_\tau$ and $\{\} \vdash \fall{x_{\tau'}}in_{\tau'}(out_{\tau'}(x_{\tau'})) = x_{\tau'}$, that are added to the global theory.

@@

In this section, we describe the design and implementation of the Supervisionary kernel.
We use a semi-formal approach, adopting a dialect of Separation Logic to describe the action of the kernel ABI.
We begin by introducing some foundational material that will be used in the remainder of the section.

\subsection{Preliminary material}

\begin{definition}
\label{defn.kernel.handles}
We fix a countably infinite set of \deffont{handles}.
We use $h$, $i$, $j$, and so on, to range arbitrarily over handles.
\end{definition}

\begin{definition}
\label{defn.partial.function}
Write $f : A \rightharpoonup B$ to denote a \deffont{partial function} from $A$ to $B$.
\end{definition}

We write $dom(f)$ for the set $\{ x \mid f(x) \text{ is defined} \}$.
We write $\emptyset$ for the \deffont{empty} partial function where $dom(\emptyset) = \{\}$ and write $a \mapsto b$ for the \deffont{singleton} partial function where $dom(a \mapsto b) = \{a\}$.

If $dom(f) \cap dom(g) = \{\}$ we write $f \uplus g$ for the partial function where $(f \uplus g)(x) = f(x)$ if $x \in dom(f)$, $(f \uplus g)(x) = g(x)$ if $x \in dom(g)$, and undefined otherwise.
Note that $f \uplus g$ is well-defined, owing to the assumption that $dom(f) \cap dom(g) = \{\}$.

\subsection{The kernel ABI}

The Supervisionary kernel maintains a mutable runtime state consisting of a series of mutable \deffont{heaps}.
These heaps are indexed by \deffont{kernel handles} which can be thought of as pointers to \deffont{kernel objects} stored in the heaps.
Importantly, the kernel never exposes kernel objects to untrusted code directly, but performs all manipulations on these objects itself, under instruction by untrusted code.

Note that in our implementation of Supervisionary handles are implemented as unsigned integers.
These are generated fresh by the kernel.
Client code may compare handles for equality and inequality, but the order in which handles are generated is undefined.

@@

The simplest---and most foundational---of these additional heaps is the \deffont{heap of type-formers}, which records the arities of type-formers registered with the kernel.

When untrusted code requests that a new type-former is registered with the kernel, it calls an ABI function---\texttt{TypeFormer.Handle.Register}---with a supplied arity.
In response, the kernel invents a fresh handle and associates it with the supplied arity, before returning the new handle back to untrusted code to use to refer to the type-former again in future.
Note that we can describe the action of the various ABI functions exposed by the kernel with Hoare triples, using an informal Separation Logic as the assertion language.

@@

\begin{figure}[t]
\caption{The internal structure of the kernel's type-former and type heap}
\label{fig.type.heap}
\end{figure}

@@

Separation Logic is a modal logic typically used to reason about pointer-programs and dynamic memory allocation.
In this section, we sketch how the Supervisionary kernel ABI can be formally specified using Separation Logic.

\begin{definition}
\label{defn.type.kernel.objects}
Define \deffont{type kernel objects} by the following grammar:
\begin{gather*}
\tau, \tau', \tau'' ::= \mathtt{TV}(n) \mid \mathtt{TC}(h, h_1, \ldots, h_n)
\end{gather*}
\end{definition}

Here, $\mathtt{TV}(n)$ represents a HOL type-variable with name $n$, whilst $\mathtt{TC}(h, h_1, \ldots, h_n)$ represents a HOL compound type, wherein $h$ points-to a registered type-former and $h_i$, for $1 \leq i \leq n$, point-to type arguments.

\begin{definition}
\label{defn.term.kernel.objects}
Define \deffont{term kernel objects} by the following grammar:
\begin{gather*}
r,s,t ::= \mathtt{V}(n, h) \mid \mathtt{C}(h, i) \mid \mathtt{A}(h, i) \mid \mathtt{L}(n, h, i)
\end{gather*}
\end{definition}

\noindent
Intuitively, we have:
\begin{itemize}
\item
$\mathtt{V}(n, h)$ represents a HOL term variable, with name $n$, wherein $h$ points-to the type of a variable.
\item
$\mathtt{C}(h, i)$ represents a HOL term constant, wherein $h$ points-to a registered constant, and $i$ points-to a type.
\item
$\mathtt{A}(h, i)$ represents the application of a functional term to an argument, wherein both $h$ and $i$ point-to terms.
\item
$\mathtt{L}(n, h, i)$ represents a $\lambda$-abstraction, with bound name $n$, wherein $h$ points-to a type and $i$ points-to a term.
\end{itemize}

\begin{definition}
\label{defn.theorem.kernel.objects}
Define \deffont{theorem kernel objects} by the following grammar:
\begin{gather*}
t, u, v ::= \mathtt{Thm}(h_1, \ldots, h_n, h)
\end{gather*}
\end{definition}

Here, $\mathtt{Thm}(h_1, \ldots, h_n, h)$ represents a HOL theorem, wherein $h_i$, for $1 \leq i \leq n$, and $h$ point-to terms---the hypotheses and the conclusions of the theorem, respectively.

\begin{definition}
\label{defn.heaps}
We define the type-former heap 
\end{definition}
